---
title: "Prepare table 2 and figure 2"
output: 
  html_document:
    toc: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading packages}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(tableone)
library(memisc)
library(sjlabelled)
library(mi)
library(boot)
library(glmnet)
library(UpSetR)
library(naniar)
library(RColorBrewer)
library(svyVGAM)
library(jtools)
library(reshape2)
```

# Read in the data

```{r}
folder_path <- "/Users/tamarsofer/Library/CloudStorage/OneDrive-BethIsraelLaheyHealth/Ongoing_papers/2022_gendered_indices"
data_with_sleep <- read_csv(file.path(folder_path, "Data/sol_sofer_sorajja_joon_covariates_20220810.csv"))

# load the prepared dataset
dat <- readRDS(file.path(folder_path, "Data/Data_with_miss.Rds"))

```

# Prepare for table 2

```{r}
# survey object
survey_base <- svydesign(id = ~PSU_ID,
                          strata = ~STRAT,
                          weights = ~WEIGHT_FINAL_NORM_OVERALL,
                          nest = TRUE,
                          data = dat)
sleepDesign <- subset(survey_base, complete.cases(dat))

# baseline covariates to adjust for
adj_vars <- c("Age", "Center", "Background")

# multi-level variables to be analyzed in multinomial regression
multi_vars <- c("Marital_status", 
                "Occupation",
                "Income_level",
                "Employment_status", 
                "Education", 
                "Years_in_US")
# not multi-level ones:
single_vars  <- c("Language_pref",
                 "Current_Health_insurance",
                 "Language_acculturation",
                  "Social_acculturation",
                  "Ethnic_identity_score", 
                  "STAI10",
                  "CESD")

cont_vars  <- c( "Language_acculturation",
                  "Social_acculturation",
                 "Ethnic_identity_score", 
                  "STAI10",
                  "CESD")

binary_vars  <- c("Language_pref",
                 "Current_Health_insurance")


```

# functions to extract summarized results from analysis, when gender is exposure
```{r}
extract_mult_out <- function(mod, exposure, out_name, level_names, round_digit = 2){
  est <- summary(mod)$coeftable
  confint <- confint(mod)
  
  inds <- grep(exposure, rownames(est))
  df_out <- data.frame(outcome = paste0(out_name,": ", level_names), 
                       est = round(exp(est[inds, "Coef"]),round_digit), 
                       CI = paste0("(", round(exp(confint[inds, 1]), round_digit), ",",
                                   round(exp(confint[inds, 2]), round_digit), ")"),
                       pval = formatC(est[inds, "p"], digits = round_digit, format = "E"))
  return(df_out)
}


extract_one_out <- function(mod, exposure, out_name, exponentiate = TRUE, round_digit = 2){
   est <- summary(mod)$coef
   confint <- confint(mod)
  
  ind <- grep(exposure, rownames(est))
  
  if (exponentiate){
      df_out <- data.frame(outcome = out_name, 
                       est = round(exp(est[ind, "Estimate"]),round_digit), 
                       CI = paste0("(", round(exp(confint[ind, 1]), round_digit), ",",
                                   round(exp(confint[ind, 2]), round_digit), ")"),
                       pval = formatC(est[ind, "Pr(>|t|)"], digits = round_digit, format = "E"))
  } else{
     df_out <- data.frame(outcome = out_name, 
                       est = round(est[ind, "Estimate"],round_digit), 
                       CI = paste0("(", round(confint[ind, 1], round_digit), ",",
                                   round(confint[ind, 2], round_digit), ")"),
                       pval = formatC(est[ind, "Pr(>|t|)"], digits = round_digit, format = "E"))
  }

  return(df_out)
}


```

# Prepare a table of results: gender is exposure

```{r}
all_out <- vector(mode = "list", length = length(multi_vars) + length(single_vars))
ind <- 1

for (var in multi_vars){
  vglm_mod <- svy_vglm(formula = as.formula(paste0(var, "~Gender+Age+Center+Background")), design = sleepDesign, family = multinomial(refLevel = 1), drop.unused.levels = TRUE)
  
  all_out[[ind]] <- extract_mult_out(vglm_mod, 
                                     exposure = "Gender", 
                                     out_name  = var, 
                                     level_names  = levels(dat[[var]])[-1])
  ind <- ind + 1
}


for (var in cont_vars){
  mod <- svyglm(formula = as.formula(paste0(var, "~Gender+Age+Center+Background")), design = sleepDesign, family = quasipoisson())
  
  all_out[[ind]] <- extract_one_out(mod, 
                                     exposure = "Gender", 
                                     out_name  = var, 
                                    exponentiate = FALSE)
  ind <- ind + 1
}

for (var in binary_vars){
  mod <- svyglm(formula = as.formula(paste0(var, "~Gender+Age+Center+Background")), design = sleepDesign, family = quasibinomial())
  
  all_out[[ind]] <- extract_one_out(mod, 
                                     exposure = "Gender", 
                                     out_name  = var, 
                                    exponentiate= TRUE)
  ind <- ind + 1
}


gender_tab <- do.call(rbind, all_out)
rownames(gender_tab) <- paste0("r", 1:nrow(gender_tab))
print(gender_tab)

write.csv(gender_tab, file = file.path(folder_path, "Results/All_vars_assoc_gender.csv"))

```

# functions to extract summarized results when insomnia is outcome


```{r}
extract_mult_exp <- function(mod, exposure, level_names, round_digit = 2){
  est <- summary(mod)$coef
  confint <- confint(mod)
  
  inds <- grep(exposure, rownames(est))
  df_out <- data.frame(exposure = paste0(exposure,": ", level_names), 
                       est = round(exp(est[inds, "Estimate"]),round_digit), 
                        CI = paste0("(", round(exp(confint[inds, 1]), round_digit), ",",
                                   round(exp(confint[inds, 2]), round_digit), ")"),
                       pval = formatC(est[inds, "Pr(>|t|)"], digits = round_digit, format = "E"))
  return(df_out)
}


extract_one_exp <- function(mod, exposure, round_digit = 2){
   est <- summary(mod)$coef
   confint <- confint(mod)
  
  ind <- grep(exposure, rownames(est))
  
  df_out <- data.frame(exposure = exposure, 
                       est = round(exp(est[ind, "Estimate"]),round_digit),
                        CI = paste0("(", round(exp(confint[ind, 1]), round_digit), ",",
                                   round(exp(confint[ind, 2]), round_digit), ")"),
                       pval = formatC(est[ind, "Pr(>|t|)"], digits = round_digit, format = "E"))
 

  return(df_out)
}


```

# Prepare a table of results: insomnia is outcome, gender unadjusted

```{r}
all_out <- vector(mode = "list", length = length(multi_vars) + length(single_vars))
ind <- 1

for (var in multi_vars){
   mod <- svyglm(formula = as.formula(paste0("Insomnia~Age+Center+Background+", var)), 
                 design = sleepDesign, family = quasibinomial())
   

  all_out[[ind]] <- extract_mult_exp(mod, 
                                     exposure = var, 
                                     level_names  = levels(dat[[var]])[-1])
  ind <- ind + 1
}


for (var in c(cont_vars, binary_vars)){
  mod <- svyglm(formula = as.formula(paste0("Insomnia~Age+Center+Background+", var)), 
                design = sleepDesign, family = quasibinomial())
  
  all_out[[ind]] <- extract_one_exp(mod, 
                                     exposure = var)
  ind <- ind + 1
}


insomnia_no_gender_tab <- do.call(rbind, all_out)
rownames(insomnia_no_gender_tab) <- paste0("r", 1:nrow(gender_tab))
print(insomnia_no_gender_tab)

write.csv(insomnia_no_gender_tab, file = file.path(folder_path, "Results/All_vars_assoc_insomnia_no_gender.csv"))
```


# Prepare a table of results: insomnia is outcome, gender adjusted

```{r}
all_out <- vector(mode = "list", length = length(multi_vars) + length(single_vars))
ind <- 1

for (var in multi_vars){
   mod <- svyglm(formula = as.formula(paste0("Insomnia~Gender + Age+Center+Background+", var)), 
                 design = sleepDesign, family = quasibinomial())
   

  all_out[[ind]] <- extract_mult_exp(mod, 
                                     exposure = var, 
                                     level_names  = levels(dat[[var]])[-1])
  ind <- ind + 1
}


for (var in c(cont_vars, binary_vars)){
  mod <- svyglm(formula = as.formula(paste0("Insomnia~Gender+Age+Center+Background+", var)), 
                design = sleepDesign, family = quasibinomial())
  
  all_out[[ind]] <- extract_one_exp(mod, 
                                     exposure = var)
  ind <- ind + 1
}


insomnia_gender_adj_tab <- do.call(rbind, all_out)
rownames(insomnia_gender_adj_tab) <- paste0("r", 1:nrow(gender_tab))
print(insomnia_gender_adj_tab)
write.csv(insomnia_gender_adj_tab, file = file.path(folder_path, "Results/All_vars_assoc_insomnia_gender_adj.csv"))
```


# Heatmap visualizing associations

For the heatmap, we will consolidate results from all models. We will use the log(OR) scale.

```{r}
# our 3 DFs with results are
# gender_tab
# insomnia_no_gender_tab
# insomnia_gender_adj_tab

heatmap_dat <- data.frame(Gender = gender_tab$est, 
                          Insomnia_gender_unadjusted = insomnia_no_gender_tab$est,
                          Insomnia_gender_adjusted = insomnia_gender_adj_tab$est)
rownames(heatmap_dat) <- gender_tab$outcome
# check:
all(gender_tab$outcome == insomnia_gender_adj_tab$exposure) # TRUE

heatmap_dat$Insomnia_gender_unadjusted <- log(heatmap_dat$Insomnia_gender_unadjusted)
heatmap_dat$Insomnia_gender_adjusted <- log(heatmap_dat$Insomnia_gender_adjusted)
inds_transform_gender <- which(!is.element(rownames(heatmap_dat), cont_vars))
heatmap_dat$Gender[inds_transform_gender] <- log(heatmap_dat$Gender[inds_transform_gender] )

# change some of variables names in the heatmap
rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Language_acculturation")] <- "Short language acculturation scale"
rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Social_acculturation")] <- "Short social acculturation scale"
rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Ethnic_identity_score")] <- "Ethnic identification score"
rownames(heatmap_dat)[which(rownames(heatmap_dat) == "CESD")] <- "CESD9"

rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Current_Health_insurance")] <- "Has current health insurance"

rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Language_pref")] <- "Language preference: English"

rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Occupation: Professional/technical, administrative/executive, or office staff")] <- "Occupation: Professional/technical"

rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Employment_status: Employed part-time(<=35 hours/week)")] <- "Employment: part-time(<=35 hours/week)"

rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Employment_status: Employed full-time(>35 hours/week)")] <- "Employment: full-time(>35 hours/week)"

rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Marital_status: Married or living with a partner")] <- "Married or living with a partner"
rownames(heatmap_dat)[which(rownames(heatmap_dat) == "Marital_status: Separated,divorced,or widow(er)")] <- "Separated,divorced,or widow(er)"


heatmap_dat$y <- rownames(heatmap_dat)

## need to set the levels according to the ordered desired in the figure!
## doing it manually...
heatmap_dat$y <- factor(heatmap_dat$y, levels = heatmap_dat$y[rev(c(1:2,  #marital status
                                                                7:10, # income
                                                                11:12, # employment
                                                                3:6, # occupation
                                                                13:14, # education
                                                                23, # health insurance 
                                                                15:16, # years in US
                                                                22, # language preference
                                                                17:19, # acculturation scales
                                                                20:21))]) # psychological

heatmap_long <- pivot_longer(heatmap_dat, 
                             cols = c("Gender", "Insomnia_gender_unadjusted", "Insomnia_gender_adjusted"),
                             names_to = "x",
                             values_to = "value")

heatmap_long$x[which(heatmap_long$x == "Insomnia_gender_adjusted")] <- "Insomnia: sex adjusted" # "Insomnia:\n sex\n adjusted"
heatmap_long$x[which(heatmap_long$x == "Insomnia_gender_unadjusted")] <- "Insomnia: unadjusted to sex" # "Insomnia:\n unadjusted\n to sex"
heatmap_long$x[which(heatmap_long$x == "Gender")] <- "Male sex" #"Male\n sex"

ggplot(heatmap_long, aes(x = x, y = y)) + 
    geom_tile(aes(fill = value)) + 
      ggtitle("Adjusted associations with sex and insomnia") +   
      scale_fill_gradient2(low = "navy", mid = "white", high = "red") + 
      ylab("") + xlab("") + 
      theme(axis.text.x = element_text( hjust= 0, size = 12), 
            axis.text.y = element_text(size = 12)) + 
          scale_x_discrete(labels = function(x) str_wrap(x, width = 7))

ggsave(file = file.path(folder_path, "Results/Heatmap_sex_insomnia_assoc.pdf"), width = 7, height = 6)

```






```{r}
sessionInfo()
```