---
title: "Train gendered indices over the complete data"
output: 
  html_document:
    toc: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading packages}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(memisc)
library(sjlabelled)
library(boot)
library(glmnet)
library(UpSetR)
library(naniar)
library(RColorBrewer)
library(jtools)
library(reshape2)
```

# Load the dataset

```{r}
folder_path <- "/Users/tamarsofer/Library/CloudStorage/OneDrive-BethIsraelLaheyHealth/Ongoing_papers/2022_gendered_indices"
data_with_sleep <- read_csv(file.path(folder_path, "Data/sol_sofer_sorajja_joon_covariates_20220810.csv"))

# load the prepared dataset
dat <- readRDS(file.path(folder_path, "Data/Data_with_miss.Rds"))
```

# Train the two indices
Primary genedered index GISE: does not include psychological variables. Secondary gendered index GIPSE: includes also trait anxiety scale and trait depression scale (without the sleep-related item).



```{r}
# Use only complete data in primary analysis
dat_cc <- dat[complete.cases(dat),]

# define the model matrix for the GISE
X_gise <- model.matrix(Sex ~
                    Marital_status +
                    Income_level + 
                    Employment_status + 
                    Occupation + 
                    Education + 
                    Current_Health_insurance + 
                    Language_pref + 
                    Language_acculturation + 
                    Social_acculturation + 
                    Ethnic_identity_score + 
                    Years_in_US, dat_cc)[,-1]



X_gipse <- model.matrix(Sex ~
                    Marital_status +
                    Income_level + 
                    Employment_status + 
                    Occupation + 
                    Education + 
                    Current_Health_insurance + 
                    Language_pref + 
                    Language_acculturation + 
                    Social_acculturation + 
                    Ethnic_identity_score + 
                    Years_in_US + 
                    STAI10 + 
                    CESD9, dat_cc)[,-1]

# GISE
set.seed(929)
cv_lasso_gise <- cv.glmnet(X_gise, dat_cc$Sex, alpha = 1, family = "binomial", nfolds = 5)

# use the lambda that resulted in minimum CVM to refit the model

set.seed(929)
lasso_model_gise <- glmnet(X_gise, dat_cc$Sex, family = "binomial", alpha = 1, lambda = cv_lasso_gise$lambda.min)
gise <- lasso_model_gise %>% predict(newx = X_gise)

# GIPSE
cv_lasso_gipse <- cv.glmnet(X_gipse, dat_cc$Sex, alpha = 1, family = "binomial", nfolds = 5)

# use the lambda that resulted in minimum CVM to refit the model

set.seed(929)
lasso_model_gipse <- glmnet(X_gipse, dat_cc$Sex, family = "binomial", alpha = 1, lambda = cv_lasso_gipse$lambda.min)
gipse <- lasso_model_gipse %>% predict(newx = X_gipse)

## add to the dataset
dat_cc$GISE <- gise
dat_cc$GIPSE  <- gipse

# save the dataset
saveRDS(dat_cc, file.path(folder_path, "Data/Data_complete_case_with_indices.Rds"))
```

# Summarize the results
## Create table 3: coefficients of the indices

```{r}
# check
all(rownames(lasso_model_gipse$beta)[1:(ncol(X_gipse)-2)] == rownames(lasso_model_gise$beta)) # TRUE

coef_gise <- c(lasso_model_gise$beta[,1], rep(NA, 2))
coef_gipse <- c(lasso_model_gipse$beta[,1])
table3 <- data.frame(Variable = colnames(X_gipse), 
                     Coef_GISE = round(coef_gise,2 ),
                     Coef_GIPSE = round(coef_gipse,2 ))

write.csv(table3, file = file.path(folder_path, "Results/Indices_coefs.csv"))
```

## Creat figure 2: index distributions by sex
We are not standardizing them at this point, because it does not matter. 

```{r}
dat_cc$Sex <- dat_cc$Sex
for_plot <- pivot_longer(dat_cc, 
                         cols = c("GISE", "GIPSE"), 
                         names_to = "Index_type", 
                         values_to = "Index")

p <- ggplot(for_plot, aes(x = Index)) + 
    geom_histogram(aes(fill = Sex), alpha = 0.6, position = "identity") +
    ggtitle("Histogram of indices stratitifed by sex") + 
    theme_bw() + 
    scale_fill_brewer(palette="Dark2") + 
    xlab("Index value") + ylab("Count") 

p + facet_wrap(~Index_type)

ggsave(file = file.path(folder_path, "Results/Indices_histograms.pdf"), 
       width = 6, height = 2.7)
```

```{r}
sessionInfo()
```