---
title: "Secondary analysis: independent data split to validate gender/sex association"
output: 
  html_document:
    toc: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading packages}
library(tidyverse)
library(survey)
library(plyr)
library(dplyr)
library(factoextra)
library(labelled)
library(memisc)
library(sjlabelled)
library(boot)
library(glmnet)
library(UpSetR)
library(naniar)
library(RColorBrewer)
library(jtools)
library(reshape2)
```
Here we create gendered indices using ~70% of the data and then evaluate them on the remaining 30% of the data, to check whether the trained indices are still distribute differently between males and females in an independent dataset.

When we split the data, we split by PSU, because individuals within the same PSU may be somewhat correlated to each other. 


# Load the dataset

```{r}
folder_path <- "/Users/tamarsofer/Library/CloudStorage/OneDrive-BethIsraelLaheyHealth/Ongoing_papers/2022_gender_measure"
data_with_sleep <- read_csv(file.path(folder_path, "Data/sol_sofer_sorajja_joon_covariates_20220810.csv"))

# load the prepared dataset
dat <- readRDS(file.path(folder_path, "Data/Data_with_miss.Rds"))
```

# Train the two indices
Primary genedered index: does not include psychological variables. Secondary gendered index: includes also trait anxiety scale and trait depression scale (without the sleep-related item).



```{r}
# Use only complete data in primary analysis
dat_cc <- dat[complete.cases(dat),]

# define train and test indices:
PSU_train <- sample(unique(dat_cc$PSU_ID), 
                    size = ceiling(0.7*length(unique(dat_cc$PSU_ID))))

ID_train <- dat_cc$ID[which(dat_cc$PSU_ID %in% PSU_train)]
ID_test <- setdiff(dat_cc$ID, ID_train)

inds_train <- which(dat_cc$ID %in% ID_train)
inds_test <- which(dat_cc$ID %in% ID_test)

length(ID_train)
length(ID_test)


# define the model matrix for the primary index 
X_primary <- model.matrix(Gender ~
                    Marital_status +
                    Income_level + 
                    Employment_status + 
                    Occupation + 
                    Education + 
                    Current_Health_insurance + 
                    Language_pref + 
                    Language_acculturation + 
                    Social_acculturation + 
                    Ethnic_identity_score + 
                    Years_in_US, dat_cc)[,-1]

X_secondary <- model.matrix(Gender ~
                    Marital_status +
                    Income_level + 
                    Employment_status + 
                    Occupation + 
                    Education + 
                    Current_Health_insurance + 
                    Language_pref + 
                    Language_acculturation + 
                    Social_acculturation + 
                    Ethnic_identity_score + 
                    Years_in_US + 
                    STAI10 + 
                    CESD, dat_cc)[,-1]

# primary index
set.seed(929)
cv_lasso_primary <- cv.glmnet(X_primary[inds_train,], dat_cc$Gender[inds_train], alpha = 1, family = "binomial", nfolds = 5)

# use the lambda that resulted in minimum CVM to refit the model

set.seed(929)
lasso_model_primary <- glmnet(X_primary[inds_train,], 
                              dat_cc$Gender[inds_train], 
                              family = "binomial", 
                              alpha = 1, 
                              lambda = cv_lasso_primary$lambda.min)
primary_index_test <- lasso_model_primary %>% predict(newx = X_primary[inds_test,])

# secondary index
cv_lasso_secondary <- cv.glmnet(X_secondary[inds_train,],
                                dat_cc$Gender[inds_train], 
                                alpha = 1, 
                                family = "binomial", 
                                nfolds = 5)

# use the lambda that resulted in minimum CVM to refit the model

set.seed(929)
lasso_model_secondary <- glmnet(X_secondary[inds_train,], 
                                dat_cc$Gender[inds_train], 
                                family = "binomial", 
                                alpha = 1, 
                                lambda = cv_lasso_secondary$lambda.min)
secondary_index_test <- lasso_model_secondary %>% predict(newx = X_secondary[inds_test,])

## add to the dataset
dat_test <- dat_cc[inds_test,]
dat_test$Primary_index  <- primary_index_test
dat_test$Secondary_index  <- secondary_index_test
```

# Summarize the results


## Creat supplemental figure 1: index distributions by gender
We are not standardizing them at this point, because it does not matter. 

```{r}
dat_test$Sex <- dat_test$Gender
for_plot <- pivot_longer(dat_test, 
                         cols = c("Primary_index", "Secondary_index"), 
                         names_to = "Index_type", 
                         values_to = "Index")

p <- ggplot(for_plot, aes(x = Index)) + 
    geom_histogram(aes(fill = Sex), alpha = 0.6, position = "identity") +
    ggtitle("Histogram of indices stratitifed by sex") + 
    theme_bw() + 
    scale_fill_brewer(palette="Dark2") + 
    xlab("Index value") + ylab("Count") 

p + facet_wrap(~Index_type)

ggsave(file = file.path(folder_path, "Results/Indices_histograms_test_set.pdf"), 
       width = 6, height = 2.7)
```

```{r}
sessionInfo()
```